---
format: 
  html:
    code-fold: true
knitr:
  opts_chunk:
    ft.htmlscroll: false
---
# Negative Binomial Hurdle Models with Zero-Heavy Count Data {.unnumbered}
**Modelling Alcohol and Other Drug Use, Psychological Distress, and Occupational Functioning in Australian Public Safety Personnel**
```{r}
#| label: setup
#| include: false
pacman::p_load(
  here,
  tidyverse,
  skimr,
  gtsummary,
  flextable,
  officer,
  scales,
  rstatix,
  GGally,
  janitor,
  DataExplorer,
  sjmisc,
  vtable,
  psych,
  car,
  lmtest,
  naniar,
  GGally,
  ggpubr,
  margins,
  cowplot,
  MASS,
  pscl,
  AER,
  performance,
  apaTables,
  Hmisc,
  glmmTMB,
  margins,
  dplyr,
  SmartEDA,
  summarytools,
  ggstatsplot,
  rcompanion,
  dunn.test
)

here::i_am("posts/hurdles.qmd")
df <- readRDS(here::here("R/clean_self_assessment.RData"))
df <- df |> filter(assessment_time == "Time 1")
df <- df |> filter(gender != "Other")
df <- df |> filter(gender != "Prefer not to answer")
source(here("R/customflextable.R"))
knitr::opts_chunk$set(message = FALSE)
min_median_max <- function(data, variable, ...) {
  x <- data[[variable]]
  tibble(
    Min = min(x, na.rm = TRUE),
    Median = median(x, na.rm = TRUE),
    Max = max(x, na.rm = TRUE)
  )
}
```
::: {.callout-note}
## Working notes 
*converting analysis script into a documented notebook*
:::

## Aims
1. Describe the rates and severity of alcohol and other drug (AOD) use in a sample of Australian public safety personnel (PSP)
2. Examine associations between AOD use, psychological distress, and occupational functioning (absenteeism and presenteeism)
3. Test whether AOD use and psychological distress, independently and together, predict occupational impairment

## Key Variables
| Role | Variable(s) |
|---|---|
| Psychological distress | `k10_total`, `k10_total_interpretation` |
| Alcohol use | `alcohol_use_scores`, `alcohol_use_interpretation` |
| Other drug use | `max_other_drug_score`, `max_drug_use_interpretation` |
| Occupational functioning | `k10_11_absenteeism`, `k10_12_presenteeism` |
| Potential covariates | `gender`, `age`, `agency_years` |

## Data Pre-processing
**exclusions:**

- dropped cases with no response to `assist1_string` - non completion of self-assessment
- excluded responses prior to march 11 (pre implementation / testing)
- set implausible ages to `NA` 

**ASSIST scoring:**

- parsed multi-response items from `assist1_string` (substances _ever_ used) to binary indicators per substance 
- non-endorsement of use treated as `NA` *not* zero (need the distinction for later models as never used ≠ low risk)
- set total scores to `NA` for substances never used for same reason
- chose each participants alcohol use score, and maximum other drug use score
- calculated interpretation categories for alcohol and other drug use seperately 
  - alcohol use: <10 = low risk, 10-26 = moderate risk, >26 = high risk
  - other drug use: <4 = low risk, 4-26 = moderate risk, >26 = high risk

**K10 scoring:**

- caclulated total k10 scores using k10+ recommended scoring 
- calculated interpretation category 
  - psychological distress: <16 = low, 16-21 = moderate, 22-29 = high, >29 = very high

**derived variables:**

- binary indicators of recent AOD use (past 3 months) - alcohol and other drugs seperately 
- poly-drug use indicator
- count of substances used in past 3 months 
- maximum other drug (non-alcohol) use risk score (self-assessment personalised feedback based on this score) 
- psychological distress total and interpretation 

::: {.callout-important} 
## Decision - drop polydrug use and substance count vars
KM and I agreed that these add little analytic value. Relying on risk scores instead. Participants only get risk score if they use in last 3 months; score = 0 if they have a life time use score; score = `NA` if they never used substance 
:::

## Screening 
Before running models, i checked distributions of all continuous outcomes and predictors. 

- Shapiro-wilk tests confirm violations of normality 
- Non-parametric tests used for group comparisons (Kruskal-Wallis) that use risk categories as grouping variable 

::: {.callout-tip}
## use custom table function
format tables automatically using [custom table function](https://amxlia.github.io/posts/customflextable.html)
:::
```{r}
#| label: tbl-shapiro
#| tbl-cap: "Shapiro tests of numeric variables"

# Shapiro-Wilk for all relevant vars
options(scipen = 999)
tbl_shapiro <- df |>
  # select vars 
  dplyr::select(
    age,
    agency_years,
    alcohol_use_scores,
    max_other_drug_score,
    k10_total,
    k10_11_absenteeism,
    k10_12_presenteeism
  ) |>
  # pivot longer so variable name is one col and shapiro stat is the other 
  pivot_longer(everything(), names_to = "var", values_to = "value") |>
  group_by(var) |>
  shapiro_test(value) |>
  # rename each var for the table 
  mutate(
    var = c(
      "Age",
      "Agency Years",
      "Alcohol Use Risk",
      "Other Drug Use Risk",
      "Psychological Distress",
      "Absenteeism",
      "Presenteeism"
    ),
    variable = NULL,
    # round numbers 
    statistic = number(statistic, accuracy = .001),
    p = number(p, accuracy = .001)
  ) |>
  rename(Variable = var, Statistic = statistic) |>
  # convert to flextable 
  flextable() |> 
  my_flextable()

tbl_shapiro
```
All variables are non-normally distributed. Not surprising given clin population and skew of outcome measures 

```{r}
#| label: fig-histograms
#| fig-cap: "Histograms of absenteeism and presenteeism"
#| warning: false
# visualise absenteeism and presenteeism 
fig_histograms <- df |>
  dplyr::select(
    Absenteeism = k10_11_absenteeism,
    Presenteeism = k10_12_presenteeism
  ) |>
  pivot_longer(
    cols = c(Absenteeism, Presenteeism),
    names_to = "Variable",
    values_to = "Days"
  ) |>
  filter(!is.na(Days)) |>
  gghistogram(
    x = "Days",
    y = "count",
    xlab = "Days",
    ylab = FALSE,
    add = "median",
    facet.by = "Variable",
    binwidth = 1
  ) +
  scale_x_continuous(breaks = seq(0, 30, by = 5))
fig_histograms
```
::: {.callout-caution}
## both outcomes have strong right skew and large proportion of zeros
suggests standard linear models would be inappropriate. might be better to work with count based models?
:::

## Prevalence and severity of AOD use 
- using ASSIST risk thresholds by substance
- excluding other and smoking from substance breakdown. 


```{r}
#| label: tbl-rates
#| tbl-cap: "Rates of substance use in past three months by risk category"

# pivot assist vars to long 
# apply risk categories 
tbl_rates <- df |>
  dplyr::select(starts_with("assist_total_"), -assist_total_scores_count) |>
  pivot_longer(
    cols = starts_with("assist_total_"),
    names_to = "substance",
    values_to = "score"
  ) |>
  mutate(score = case_when(score < 1 ~ NA, TRUE ~ score)) |>
  filter(!is.na(score)) |>
  mutate(
    substance = str_remove(substance, "assist_total_") |> str_to_title()
  ) |>
  mutate(
    `Risk` = case_when(
      substance == "Alcohol" & score < 10 ~ "Low risk",
      substance == "Alcohol" &
        score >= 10 &
        substance == "Alcohol" &
        score <= 26 ~ "Moderate risk",
      substance == "Alcohol" & score > 26 ~ "High risk",
      substance != "Alcohol" & score < 4 ~ "Low risk",
      substance != "Alcohol" &
        score >= 4 &
        substance != "Alcohol" &
        score <= 26 ~ "Moderate risk",
      substance != "Alcohol" & score > 26 ~ "High risk",
      TRUE ~ "fix"
    )
  ) |>
  group_by(substance, Risk) |>
  group_by(substance, Risk) |>
  summarise(
    n = n()
  ) |>
  ungroup() |>
  group_by(substance) |>
  mutate(total = sum(n), percent = round(100 * n / sum(n), 1)) |>
  group_by(substance, Risk) |>
  ungroup() |>
  dplyr::select(substance, total, Risk, n, percent) |>
  pivot_wider(names_from = Risk, values_from = c(n, percent)) |>
  dplyr::select(
    Substance = substance,
    total,
    ends_with("Low risk"),
    ends_with("Moderate risk"),
    ends_with("High risk")
  ) |>
  mutate(
    total = paste0(total, " (", number(total / 220 * 100, accuracy = .1), "%)")
  ) |>
  filter(Substance != 'Other') |>
  filter(Substance != "Smoking") |>
  flextable() 

# change spanning headers to group n and % under each risk category 
typology <- data.frame(
  col_keys = c(
    "Substance",
    "total",
    "n_Low risk",
    "percent_Low risk",
    "n_Moderate risk",
    "percent_Moderate risk",
    "n_High risk",
    "percent_High risk"
  ),
  head = c("", "", rep("ASSIST scores", 6)),
  risk = c(
    "",
    "",
    rep("Low risk", 2),
    rep("Moderate risk", 2),
    rep("High risk", 2)
  ),
  names = c("Substance", "Total using", rep(c("n", "%"), 3))
)

# print and format rates table 
tbl_rates <- tbl_rates |>
  set_header_df(mapping = typology, key = "col_keys") |>
  merge_h(part = "header") |>
    my_flextable() |> 
    hline(
      border = fp_border(color = "black", width = 1),
      i = c(1, 2),
      part = "header"
  ) |> 
    hline_top(border = fp_border(color = "black", width = 1), part = "all")

tbl_rates 
```

**observations:**

- alcohol use = nearly everyone, meaningful spread across risk levels too 
- other substance use is less prevalenct, but does seem to be more moderate risk (though the low-risk threshold is pretty low, so even occasional use is likely to be moderate??)
- high risk drug use = uncommon, but evident 
- cannabis, amphetamines, and sedatives more skewed to moderate, rather than low risk 

::: {.callout-note} 
## modelling decision
rather than modelling each drug individually, use max other drug score as indicator. 
:::

### Summary based on alcohol risk: 
- restrict to recent alcohol users (past 3 months) 
- stratify by ASSIST risk
- compare continuous with kruskal-wallis and gender with chi-squared 

```{r}
#| label: tbl-alc
#| tbl-cap: "Sample characteristics by alcohol use risk levels"
#| warning: false

# Summary based on alcohol risk scores
tbl_alc <- df |>
  filter(recent_alcohol_user == 1) |>
  dplyr::select(
    gender,
    age,
    agency_years,
    alcohol_use_interpretation,
    k10_total,
    absenteeism = k10_11_absenteeism,
    presenteeism = k10_12_presenteeism
  ) |>
  tbl_summary(
    by = alcohol_use_interpretation,
    missing = "no",
    missing_text = "Missing",
    statistic = list(
      all_continuous() ~ c("{mean} ({sd})"),
      all_categorical() ~ "{n} ({p}%)"
    ),
    percent = "row",
    digits = list(all_continuous() ~ c(1, 1), all_categorical() ~ c(0, 1)),
    label = list(
      gender = "Gender",
      age = "Age",
      k10_total = "Psychological distress",
      absenteeism = "Absenteeism",
      presenteeism = "Presenteeism",
      agency_years = "Agency years",
      alcohol_use_interpretation = "Alcohol use risk category"
    )
  ) |>
  add_n(statistic = "{N_nonmiss}", col_label = "n") |>
  modify_header(list(
    label ~ "Characteristic",
    stat_1 ~ "Low Risk\n(n = 77)",
    stat_2 ~ "Moderate Risk\n(n = 83)",
    stat_3 ~ "High Risk\n(n = 56)"
  )) |>
  add_p(
    test = list(all_continuous() ~ "kruskal.test", gender ~ "chisq.test"),

    #test.args = all_categorical() ~ list(simulate.p.value = TRUE),
    pvalue_fun = function(x) {
      formatted <- ifelse(x < 0.001, "<.001", sprintf("%.3f", x))
      sub("^0\\.", ".", formatted)
    }
  ) |>
  add_stat_label(
    label = list(all_continuous() ~ "Mean (SD)", all_categorical() ~ "n (%)")
  ) |>
  modify_footnote(everything() ~ NA) |>
  modify_header(p.value ~ "p") |>
  gtsummary::as_flex_table() |> 
  my_flextable()
tbl_alc 
```

- higher alcohol risk is associated with greater psychological distress and higher absenteeism and presenteeism 
- age and years in the agency are higher in moderate/high risk groups .. could be related to cumulative exposure to traumatic events etc. 
- post hoc dunn tests with bonferroni correction show most differences are driven by low vs. moderate/high. moderate vs. high risk are pretty similar. 
```{r}
#| eval: false
#| label: dunn-alcohol 

# post hoc pairwise comparisons - included for completeness. 
tests <- df |> filter(recent_alcohol_user == 1)
dunn.test(tests$age, tests$alcohol_use_interpretation, kw = TRUE, method = "bonferroni")
dunn.test(tests$agency_years, tests$alcohol_use_interpretation, kw = TRUE, method = "bonferroni")
dunn.test(tests$k10_total, tests$alcohol_use_interpretation, kw = TRUE, method = "bonferroni")
dunn.test(tests$k10_11_absenteeism, tests$alcohol_use_interpretation, kw = TRUE, method = "bonferroni")
dunn.test(tests$k10_12_presenteeism, tests$alcohol_use_interpretation, kw = TRUE, method = "bonferroni")

# gender - chi-squared then pairwise 
report::report(stats::chisq.test(tests$gender, tests$alcohol_use_interpretation)) 

gender_tbl <- table(tests$gender, tests$alcohol_use_interpretation)
pairwiseNominalIndependence(gender_tbl, fisher = FALSE, gtest = FALSE, chisq = TRUE, method = "bonferroni")

prop.table(table(tests$gender, tests$alcohol_use_interpretation), margin = 1)
```
### Summary based on other drug risk: 
Same as above, but for recent other drug use

- much smaller n here (especially high risk group, n = 5); interpret w caution 
- fisher's exact test used for gender here given small cell sizes 
```{r}
#| label: tbl-drug
#| tbl-cap: "Sample characteristics by other drug use risk levels"
#| warning: false
tbl_drug <- df |>
  filter(recent_drug_user == 1) |>
  dplyr::select(
    gender,
    age,
    agency_years,
    max_drug_use_interpretation,
    k10_total,
    absenteeism = k10_11_absenteeism,
    presenteeism = k10_12_presenteeism
  ) |>
  tbl_summary(
    by = max_drug_use_interpretation,
    missing = "no",
    missing_text = "Missing",
    statistic = list(
      all_continuous() ~ c("{mean} ({sd})"),
      all_categorical() ~ "{n} ({p}%)"
    ),
    percent = "row",
    digits = list(all_continuous() ~ c(1, 1), all_categorical() ~ c(0, 1)),
    type = list(
      recent_drug_count = "continuous",
      #recent_alcohol_user = "categorical",
      recent_drug_user = "categorical",
      poly_drug_user = "categorical"
    ),
    label = list(
      gender = "Gender",
      age = "Age",
      k10_total = "Psychological distress",
      absenteeism = "Absenteeism",
      presenteeism = "Presenteeism",
      agency_years = "Agency years",
      max_drug_use_interpretation = "Maximum other drug use risk category"
    )
  ) |>
  add_n(statistic = "{N_nonmiss}", col_label = "n") |>
  modify_header(list(
    label ~ "Characteristic",
    stat_1 ~ "Low Risk\n(n = 17)",
    stat_2 ~ "Moderate Risk\n(n = 46)",
    stat_3 ~ "High Risk\n(n = 5)"
  )) |>
  add_p(
    test = list(all_continuous() ~ "kruskal.test", gender ~ "fisher.test"),

    #test.args = all_categorical() ~ list(simulate.p.value = TRUE),
    pvalue_fun = function(x) {
      formatted <- ifelse(x < 0.001, "<.001", sprintf("%.3f", x))
      sub("^0\\.", ".", formatted)
    }
  ) |>
  add_stat_label(
    label = list(all_continuous() ~ "Mean (SD)", all_categorical() ~ "n (%)")
  ) |>

  modify_footnote(everything() ~ NA) |>
  modify_header(p.value ~ "p") |>
  gtsummary::as_flex_table() |>
  my_flextable()
tbl_drug
```
- psychological distress and absenteeism seem to show a dose-reponse pattern across risk groups 
- presenteeism the same direction, but not significant, possible power issues. 
- age and agency years and gender = no differences, contrast with alcohol use results. 

dunn tests aren't really reliable here (but post hoc tests w bonferroni correction code): 
```{r}
#| eval: false
#| label: dunn-otherdrugs 

# post hoc pairwise comparisons - included for completeness. 
tests <- df |> filter(recent_drug_user == 1)
dunn.test(tests$age, tests$max_drug_use_interpretation, kw = TRUE, method = "bonferroni")
dunn.test(tests$agency_years, tests$max_drug_use_interpretation, kw = TRUE, method = "bonferroni")
dunn.test(tests$k10_total, tests$max_drug_use_interpretation, kw = TRUE, method = "bonferroni")
dunn.test(tests$k10_11_absenteeism, tests$max_drug_use_interpretation, kw = TRUE, method = "bonferroni")
dunn.test(tests$k10_12_presenteeism, tests$max_drug_use_interpretation, kw = TRUE, method = "bonferroni")
```

## Associations 
- pearson for continuous-continuous pairs (alcohol score, drug score, k10 score)
- spearman for pairs involving absenteeism/presenteeism - count variables, zero inflated, rank based more appropriate (see Figure 1)

```{r}
#| label: tbl-cors 
#| tbl-cap: "Correlation matrix of key variables"
#| message: false 
# select vars 
cor_vars <- df |>
  dplyr::select(
    alcohol_use_scores, max_other_drug_score, 
    k10_total, absenteeism = k10_11_absenteeism, presenteeism = k10_12_presenteeism
  ) 

# define variable types 
continuous <- c("k10_total", "alcohol_use_scores", "max_other_drug_score")
count <- c('presenteeism', 'absenteeism', 'recent_drug_count')

# get all correlation pairs 
cor_pairs <- combn(names(cor_vars), 2, simplify = FALSE)

# create function to get correct correlation type automatically 
get_cors <- function(x_name, y_name, data) {
  x <- data[[x_name]]
  y <- data[[y_name]]
  
  # define variable types 
  is_x_count <- x_name %in% count
  is_y_count <- y_name %in% count
  is_x_continuous <- x_name %in% continuous
  is_y_continuous <- y_name %in% continuous
  
  # define correlation method 
  method <- case_when(
    is_x_count & is_y_count ~ "spearman", # use spearman
    is_x_count & is_y_continuous ~ "spearman", # use spearman
    is_x_continuous & is_y_count ~ "spearman", # use spearman
    is_x_continuous & is_y_continuous ~ "pearson",  # use pearson 
    TRUE ~ "spearman"  # if else, use spearman 
  )
  
  cor_type <- case_when(
    is_x_continuous & is_y_continuous ~ "pearson",
    TRUE ~ "spearman"
  )
  
  # calculate correlations 
  test <- suppressWarnings(cor.test(x, y, method = method))
  
  # return correlation coefficient, p-value, method used
  list(r = unname(test$estimate), p = test$p.value, method = method, type = cor_type)
}

# compute correlations for all pairs 
cor_results <- purrr::map_dfr(cor_pairs, function(vars) {
  res <- get_cors(vars[1], vars[2], cor_vars)
  tibble(
    var1 = vars[1],
    var2 = vars[2],
    cor = res$r,
    p = number(res$p, accuracy = .001),
    method = res$method,
    type = res$type#,
    #p2 = res$p
  )
})

# make symmetric by mirroring rows
cor_long <- cor_results |>
  bind_rows(
    cor_results |>
      rename(var1 = var2, var2 = var1)
  ) |>
  mutate(
    cor = round(cor, 2),
    sig = p <= 0.05,
    stars = case_when(
      p < .001 ~ "***",
      p < .01 ~ "**", 
      p < .05 ~ "*",
      TRUE ~ ""),
    cor_ch = paste0(formatC(cor, format = "f", digits = 2), stars))
  
# make wide 
cor_matrix <- cor_long |>
  dplyr::select(var1, var2, cor_ch) |>
  pivot_wider(names_from = var2, values_from = cor_ch) |> 
  dplyr::select(var1, alcohol_use_scores, max_other_drug_score, k10_total, absenteeism, presenteeism)

# set cor labels 
cor_labels <- c(
  k10_total = "Psychological Distress",
  presenteeism = "Presenteeism",
  absenteeism = "Absenteeism",
  alcohol_use_scores = "Alcohol Use Score",
  max_other_drug_score = "Max Other Drug Score")

colnames(cor_matrix) <- c(" " = "var1", cor_labels[colnames(cor_matrix)[-1]])
cor_matrix[[1]] <- cor_labels[cor_matrix[[1]]]
cor_matrix[upper.tri(cor_matrix)] <- NA

# create ft 
cor_tbl <- flextable(cor_matrix) |> 
  my_flextable() |> 
  set_header_labels("var1" = " ")

cor_tbl
```
**observations:**
Pretty much everything is highly correlated (ranging from .24 - .52) . Strongest associations are with psychological distress (particularly with absenteeism and presenteeism), but that makes conceptual sense. might be also that they come from the k10+? Questions specifically asks "due to the symptoms described" (in the k10), so naturally they would be correlatedexcept for alcohol use and other drug use.

Weaker associations between alcohol use and other drug use which is sort of good for modelling - no collinearity
 
::: {.callout-important}
## flagging for later
Low correlation between two AOD predictors suggests multicollinearity won't be an issue in models. 

**REMEMBER**: check VIFs after fitting 
:::

## Model selection
I considered several candidate models that fit the data for absenteeism and presenteeism: 

- Poisson regression: baseline count model 
- Negative binomial regression: relaxes the mean
- Zero-inflated Poisson (ZIP): models excess zeros by latent *always zero* group
- Zero-inflated negative binomial models (ZINB): combines zero-inflation with overdispersion
- Hurdle negative binomial models: predicts first whether count is zero, then models positive counts 

**checks:**
rounded absenteeism and presenteeism to 1dp 
```{r}
#| label: prep
# round absenteeism and presenteeism 
df <- df |> mutate(
  k10_11_absenteeism = round(k10_11_absenteeism, 1),
  k10_12_presenteeism = round(k10_12_presenteeism, 1)
) |> rename(absenteeism = k10_11_absenteeism, 
             presenteeism = k10_12_presenteeism)
# select vars 
df_sub <- df |> dplyr::select(age, gender, agency_years, k10_total, rtc_3_readiness, 
                               recent_drug_user, recent_alcohol_user, recent_any_user, poly_drug_user,
                               alcohol_use_scores, max_other_drug_score, absenteeism, presenteeism)
```

**wrote function for comparing models**:
instead of fitting each model manually for every predictor, function loops through all univariate predictors for a given outcome, and compares models across all 5 candidates 

predictors:

- age 
- gender
- agency years 
- k10 total
- alcohol use 
- other drug use 

each iteration returns: 

- dispersion statistics 
- AIC and BIC 
- vuong tests 
- observed vs predicted zero counts 

saved results to file (lots of output) and its really only useful for diagnostic purposes 

```{r}
#| label: model-comparison-function
#| eval: false
predictors <- df_sub |> dplyr::select(-absenteeism, -presenteeism) |> names()

# function to run series of univariate models and check dispersion. 
compare_models <- function(var, data) {
  fmla <- as.formula(paste(outcome, "~", var))

  # print results:
  cat("\n============================\n")
  cat("Variable:", var, "\n")
  cat("============================\n\n")
  
  # fit models 
  # poisson
  pois <- glm(fmla, data = data, family = poisson())
  # negative binomial
  nb <- glm.nb(fmla, data = data)
  # zero inflated poisson
  zip <- zeroinfl(fmla, data = data, dist = "poisson")
  # zero inflated negative binomial
  zinb <- zeroinfl(fmla, data = data, dist = "negbin")
  # hurdle negative binomial 
  hurdle_nb <- hurdle(fmla, data = data, dist = "negbin", zero.dist = "binomial")
  
  # dispersion:
  disp_stat <- dispersiontest(pois)$statistic 
  cat('Dispersion (Poisson):', round(disp_stat, 2), '\n')
  
  # compare AIC / BIC
  aic_df <- AIC(pois, nb, zip, zinb, hurdle_nb)
  bic_df <- BIC(pois, nb, zip, zinb, hurdle_nb)
  
  cat("\nAIC Comparison:\n")
  print(aic_df)
  cat("\nBIC Comparison:\n")
  print(bic_df)
  
  # vuong tests
  cat("\nVuong Test: ZIP vs Poisson\n")
  print(vuong(pois, zip))
  
  cat("\nVuong Test: ZINB vs NB\n")
  print(vuong(nb, zinb))
  
  cat("\nVuong Test: Hurdle vs NB\n")
  print(vuong(nb, hurdle_nb))

# get predicted vs observed zeros 
obs_zeros <- sum(data[[outcome]] == 0)
pred_zeros <- list(
  Poisson = sum(predict(pois, type = "response") < 1e-6),
  NB = sum(predict(nb, type = "response") < 1e-6),
  ZIP = sum(predict(zip, type = "response") < 1e-6),
  ZINB = sum(predict(zinb, type = "response") < 1e-6),
  Hurdle = sum(predict(hurdle_nb, type = "response") < 1e-6)
)

cat("\nObserved Zeros:", obs_zeros, "\n")
cat("Predicted Zeros:\n")
print(pred_zeros)
}

# run and save externally 
outcome <- "absenteeism"
predictors <- df_sub |> dplyr::select(-presenteeism, -absenteeism) |> names()

sink(here("R/model_comparisons_absenteeism.txt"))
lapply(predictors, compare_models, data = df_sub |> filter(!is.na(absenteeism)))
sink()

outcome <- "presenteeism"
sink(here("R/model_comparisons_presenteeism.txt"))
lapply(predictors, compare_models, data = df_sub |> filter(!is.na(presenteeism)))
sink()
```
- overdispersion confirmed across all predictors - poisson dispersion consistently >1 (not a good candidate)
- zero-inflated and hurdle models outperformed NB on AIC/BIC and predicted zeros better 
- vuong tests favoured hurdle and zero-inflated, though difference between ZINB and hurdle NB was small 


fit statistics between ZINB and hurdle NB were similar enough. MS and I discussed this, and we think zero-inflation models were too rigid for our outcomes. 

zero inflated models assume 2 groups: ppl structurally **always zero** (i.e., will never ever have any absenteeism/presenteeism) and people who are at risk. for occupational outcomes in this, employed sample, not the best assumption, anyone can have a day off or a bad day at work. 

hurdle models seem more appropriate, because it models the likelihood of someone crossing into abscence/not, and then a seperate model to determine how many days are lost if they do.... this fits our research question better... 

- part 1 (logit): what predicts whether someone reports absenteeism or presenteeism? 
- part 2 (count): among those who do report, what predicts how many days they do so for? 

::: {.callout-note}
## Decision: Hurdle negative binomial models 
They account for overdispersion, handle excess zeros in two stages, and align with our aims
:::


## Modelling  
- converted missing AOD scores to zero
- missing score means they didnt use it in the past 3 months, so zero is appriopriate for modelling rather than excluding the cases 

```{r}
#| label: model-prep 
df_sub <- df |>
  dplyr::select(
    age, gender, agency_years,
    alcohol_use_scores, max_other_drug_score,
    k10_total, absenteeism, presenteeism
  ) |>
  mutate(
    # NA = did not use in past 3 months; treat as zero risk for modelling
    alcohol_use_scores = tidyr::replace_na(alcohol_use_scores, 0),
    max_other_drug_score = tidyr::replace_na(max_other_drug_score, 0)
  )


  # variable labels used across all model tables:
  var_labels <- c(
  agency_years = "Years in agency",
  age = "Age",
  gender = "Gender",
  alcohol_use_scores = "Alcohol use risk",
  max_other_drug_score = "Maximum other drug use risk",
  k10_total = "Psychological distress",
  `max_other_drug_score:k10_total` = "Other drug use risk x Psychological distress"
)

```

### Univariate models - Absenteeism 
each predictor examined as univariate effect before building multivariable model to:

- see variable-specific effects before adding covariates 
- identify which predictors were worth using 
- check that hurdle structure worked as expected 

each univariate model produces two sets of estimates: odds ratio (OR) from the logit model ("any absenteeism"), and an incidence rate ratio (IRR) from the negative binomial part ("how many days in those with at least 1?")

wrote a function to loop this across predictors and stack results in a table (probably only 4 supp materials in publication )

```{r}
#| label: fnction-uv
#| warning: false 
run_uv_hurdle <- function(outcome, predictors, 
                          data, exponentiate = T) {
  
  # build both models seperately 
  model_tables <- map(predictors, ~ {
  fmla <- as.formula(paste0(outcome, " ~ ", .x, " | ", .x))
  model <- hurdle(fmla, data = data, dist = "negbin", zero.dist = "binomial")
    
  # zero part 
  zero_tbl <- tbl_regression(
    model,
    component = "zero_inflated",
    exponentiate = TRUE,
    estimate_fun = function(x) formatC(x, digits = 2, format = "f"),
    pvalue_fun = function(x) {
      p <- style_pvalue(x, digits = 3)
      # remove leading zero from both "0.03" and "<0.001"
      gsub("(?<=<)0\\.|^0\\.", ".", p, perl = TRUE)
    }, 
    tidy_fun = broom.helpers::tidy_zeroinfl
  ) |> 
    modify_header(label = "Variable", estimate = "OR") |> 
    modify_table_body(~ .x |>
                        mutate(
                          label = ifelse(
                            row_type == "label" & variable %in% names(var_labels),
                            var_labels[variable],
                            label
                          ),
                          ci = ifelse(
                            !is.na(estimate) & (conf.high - conf.low > 50),
                            ">50",
                            ci
                          )
                        )
    ) |> 
    add_n() 
  
  # NB count part
  count_tbl <- tbl_regression(
    model,
    component = "conditional",
    exponentiate = T,
    estimate_fun = function(x) formatC(x, digits = 2, format = "f"),
    pvalue_fun = function(x) {
      p <- style_pvalue(x, digits = 3)
      # remove leading zero from both "0.03" and "<0.001"
      gsub("(?<=<)0\\.|^0\\.", ".", p, perl = TRUE)
    },
    
    tidy_fun = broom.helpers::tidy_zeroinfl
  ) |> 
    modify_header(label = "Variable", estimate = "IRR") |> 
    modify_table_body(~ .x |>
                        mutate(
                          label = ifelse(
                            row_type == "label" & variable %in% names(var_labels),
                            var_labels[variable],
                            label
                          ),
                          ci = ifelse(
                            !is.na(estimate) & (conf.high - conf.low > 50),
                            ">50",
                            ci
                          )
                        )) 
  
  list(zero_tbl = zero_tbl, count_tbl = count_tbl)
  })

  # separate and flatten count and zero tables
  zero_tbls  <- map(model_tables, "zero_tbl")
  count_tbls <- map(model_tables, "count_tbl")
  
  tbl_merge(
    tbls = list(
      tbl_stack(zero_tbls),
      tbl_stack(count_tbls) 
      ),
    tab_spanner = c("Hurdle model (logit)", "Count model (NB)")
  )
}
```

then i applied this first to absenteeism
```{r}
#| label: tbl-abs-uv 
#| tbl-cap: "Univariable hurdle negative binomial regressions predicting absenteeism"
#| warning: false

outcome <- "absenteeism"
predictors <- df_sub |> dplyr::select(-presenteeism, -absenteeism) |> names()

tbl_abs_uv <- suppressMessages(
  run_uv_hurdle(outcome, predictors, data = df_sub, exponentiate = TRUE)
) 

tbl_abs_uv <- tbl_abs_uv |> 
  modify_table_body(~ .x |>
    mutate(
      estimate_1 = number(estimate_1, accuracy = .01),
      estimate_1 = if_else(is.na(estimate_1), "—", estimate_1),
      estimate_2 = number(estimate_2, accuracy = .01),
      estimate_2 = if_else(is.na(estimate_2), "—", estimate_2),
      label = case_when(label == "0" ~ "No", label == "1" ~ "Yes", TRUE ~ label),
      groupname_col_1 = NULL,
      groupname_col_2 = NULL
    )
  ) |>
  as_flex_table() |>
  bold(part = "all", bold = FALSE) |> 
  delete_part("footer") |> 
  my_flextable()


tbl_abs_uv
```